---
title: "Final EDA"
author: "Team 4"
date: "5/20/2022"
output: html_document
---

```{r setup, include=FALSE}
library(fpp)
library(tseries)
library(ggplot2)
library(forecast)
library(urca)
library(TSA)
library(vars)
library(shape)

library(tidyverse)
library(lubridate)
library(clifro)
library(MASS)
library(stringr)
library(zoo)
library(viridis)
library(broom)
library(plotly)
library(knitr)
library(dplyr)
library(aTSA)
```

## Load in data 

Air temperature (C), Wind Speed (m/s), Wind Direction (degrees from; 0=from N, 90=from E, etc.)

```{r}
#Read in the file list from folder
filelist <- list.files("2011-2021", full.names = TRUE)
filelist <- sort(filelist)
filelist
```


```{r}
#Initialize the first file for appending
r <- read.csv(filelist[1], sep="", header = TRUE)
head(r, 10)
```

```{r}
#Loop to read in all files 
for (i in 2:length(filelist)){
  ri <- suppressMessages(read.csv(filelist[i], sep=""))
  r <- rbind(r, ri)
} 
```

```{r}
df <- r
df$Date = seq(as.Date("2011-01-01"),as.Date("2011-01-01") + dim(r)[1] - 1,by = 1)
df$Month <- as.numeric(format(df$Date,'%m'))
df$Year <- as.numeric(format(df$Date,'%Y'))
tail(df)
```
## Visualization 

### Wind Speed (m/s)

```{r}
#As TS
ggplot(df, aes(x = DOY, y = WS, color = Month)) +
  geom_line()
#Distribution of all 
hist(df$WS, main = 'Histogram of Wind Speed (m/s)', xlab = 'Wind Speed (m/s)')

#Box plot by Month
boxplot(WS~Month,df, ylab = 'Wind Speed (m/s)')
```
### Wind Direction

```{r}
#WD visualization
windrose(speed = df$WS,
                 direction =df$WD,
                 speed_cuts = seq(0,25,5),
                 ggtheme='minimal')
```

```{r}
#As TS
ggplot(df, aes(x = DOY, y = WD, color = Month)) +
  geom_line()
#General Distribution
hist(df$WD, main = 'Histogram of Wind Direction', xlab = 'Wind Direction')
#Box plot by Month
boxplot(WD~Month,df, ylab = 'Wind Direction')
```

### Air Tempreture
```{r}
#As TS
ggplot(df, aes(x = DOY, y = AT, color = Month)) +
  geom_line()
#General Distribution
hist(df$AT, main = 'Histogram of Air Tempreture', xlab = 'Air Tempreture')
#Box plot by Month
boxplot(AT~Month,df, ylab = 'Air Tempreture')
```

We see obvious outliers from the line plot and histogram on 2019 May. 

```{r}
#qplots
qplot(WD,WS,data=df,color=Month,geom = c("point","smooth"),facets = .~Month)
qplot(AT,WS,data=df,color=Month,geom = c("point","smooth"),facets = .~Month)
```

```{r}
cor(df$WS,df$AT)
```


```{r}
cor(df$WS,df$WD)
```

A value of 0.21, -0.22 shows there is a positive correlation between two variables, but it is weak and likely unimportant.

### Data Aggregation 

We wanted to explore wind speed change monthly, hence we aggregate our data into monthly averages.


```{r}
df.agg <- df %>% group_by(Year, Month) %>% summarize(m_ws = mean(WS), m_at = mean(AT), m_wd = mean(WD))
```


```{r}
#Rounding 
df.agg$m_ws <- round(df.agg$m_ws, digits = 5)
df.agg$m_at <- round(df.agg$m_at, digits = 5)
df.agg$m_wd <- round(df.agg$m_wd, digits = 5)
```

```{r}
#Reset row index
df.agg$Order <- 1:nrow(df.agg)
df.agg
```

## Visualization of Aggregated Data

### Wind Speed (m/s)

```{r}
#As TS
ggplot(df.agg, aes(x = Order, y = m_ws, color = Month)) +
  geom_line()
#Distribution of all 
hist(df.agg$m_ws, main = 'Histogram of Wind Speed (m/s)', xlab = 'Wind Speed (m/s)')

#Box plot by Month
boxplot(m_ws~Month,df.agg, ylab = 'Wind Speed (m/s)')
```
### Wind Direction


```{r}
windrose(speed = df.agg$m_ws,
                 direction =df.agg$m_wd,
                 speed_cuts = seq(0,10,2),
                 ggtheme='minimal')
```


```{r}
#As TS
ggplot(df.agg, aes(x = Order, y = m_wd, color = Month)) +
  geom_line()
#General Distribution
hist(df.agg$m_wd, main = 'Histogram of Wind Direction', xlab = 'Wind Direction')
#Box plot by Month
boxplot(m_wd~Month,df.agg, ylab = 'Wind Direction')
```
### Air Tempreture 
```{r}
#As TS
ggplot(df.agg, aes(x = Order, y = m_at, color = Month)) +
  geom_line()
#General Distribution
hist(df.agg$m_at, main = 'Histogram of Air Tempreture', xlab = 'Wind Tempreture')
#Box plot by Month
boxplot(m_at~Month,df.agg, ylab = 'Wind Tempreture')
```
We see an extreme outlier in our air temperature dataset. 

```{r}
#Getting outliers at day to day level 
at.outliers <- boxplot(df.agg$m_at, plot=FALSE)$out
df.agg[which(df.agg$m_at %in% at.outliers),]
#which(demo.data$score > quantile(demo.data$score)[4] + 1.5*IQR(demo.data$score)
```
We see the record from 2019 May that has an unusual air temperature at -47.25484.

```{r}
hist(df[which(df$Year == 2019 & df$Month == 5),'AT'], xlab = 'Air Temperature', main= 'Daily Air Temperature from 2019 May')
```


We see an unusual pattern of distribution which have much data points at -99 air temperature. We consider them corrupted data and hence replace the monthly air temperature from 2019 May with the average of May air temperature from other years.

```{r}
#New temp as the average of May air temperature from other years
new_at <- mean(unlist(df.agg[which(df.agg$Month == 5 & df.agg$Year != 2019),'m_at']))
df.agg[which(df.agg$m_at %in% at.outliers),'m_at'] <- new_at
df.agg$m_at
```

Checking the distribution of air temperature after taking out the outlier.

```{r}
#As TS
ggplot(df.agg, aes(x = Order, y = m_at, color = Month)) +
  geom_line()
#General Distribution
hist(df.agg$m_at, main = 'Histogram of Air Tempreture', xlab = 'Wind Tempreture')
#Box plot by Month
boxplot(m_at~Month,df.agg, ylab = 'Wind Tempreture')
```
We see the outlier is removed while maintaining the original shape of distribution.


```{r}
#write.csv(r, 'all_days.csv')
write.csv(df.agg, 'agg.csv')
```
